Spatial proteomics datasets
============================

## The `pRolocdata` package

`pRolocdata` is a [Bioconductor](http://bioconductor.org/)
[experiment package](http://bioconductor.org/packages/release/BiocViews.html#___ExperimentData)
([release](http://bioconductor.org/packages/release/data/experiment/html/pRolocdata.html)
and
[devel](http://bioconductor.org/packages/devel/data/experiment/html/pRolocdata.html)
pages) that collects published (mainly, although some unpublished
datasets are also available) mass spectrometry-based spatial/organelle
and protein-complex dataset. The data are distributed as `MSnSet`
instances (see the
[`MSnbase`](http://www.bioconductor.org/packages/release/bioc/html/MSnbase.html)
for details) and are used throughout the
[`pRoloc`](http://bioconductor.org/packages/release/data/experiment/html/pRolocdata.html)
and
[`pRolocGUI`](http://bioconductor.org/packages/devel/bioc/html/pRolocGUI.html)
software for spatial proteomics data analysis and visualisation.

**Current build status**:

[![Build Status](https://travis-ci.org/lgatto/pRolocdata.svg?branch=master)](https://travis-ci.org/lgatto/pRolocdata)

### Installation

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library("knitr")
library("pRolocdata")
x <- data.frame(pRolocdata()$results[, -(1:2)])
colnames(x) <- c("Data", "Description")
```

```{r, eval=FALSE}
source("http://bioconductor.org/biocLite.R")
biocLite("pRolocdata")
```

Once installed, the package needs to be loaded

```{r, eval=FALSE}
library("pRolocdata")
```

### Available datasets

Currently, there are `r nrow(x)` datasets available in
`pRolocdata`. Use the `pRolocdata()` function to obtain a list of data
names and their description.

```{r, eval=FALSE}
pRolocdata()
```

```{r, echo=FALSE, results='asis'}
kable(x, format = "markdown")
```

### Loading data

Data is loaded into the `R` session using the `load` function; for
instance, to get the data from
[Dunkley et al (2006)](http://www.pnas.org/content/103/17/6518.abstract),
one would type

```{r}
data(dunkley2006)
```

To get more information about a given dataset, see its manual page

```{r, eval=FALSE}
?dunkley2006
## or
help("dunkley2006")
```

## The datasets

Each data object in `pRolocdata` is available as an `MSnSet`
instance. The instances contain the actual quantitative data, sample
and features annotations (see `pData` and `fData`
respectively). Additional MIAPE data
[[1](https://en.wikipedia.org/wiki/Minimum_Information_About_a_Proteomics_Experiment),
[2](http://www.nature.com/nbt/journal/v25/n8/abs/nbt1329.html)]
experimental data is available in the `experimentData` slot, as
described in section *Required metadata* below.

The source of these data is generally one or several text-based
spreadsheet (`csv`, `tsv`) produced by a third-party
application. These original files are often distributed as
supplementary material to the research paper and used to generate the
`R` objects. These source spreadsheets are available in the package's
`inst/extdata` directory. The `R` script files, that read the
spreadsheets and create the `R` data is distributed in the
`inst/scripts` directory.

### Suggested metadata

Additional metadata is available with the `pRolocmetadata()` function
as detailed below.

#### Species
Documented in `experimentData(.)@samples$species`

#### Tissue

Documented in `experimentData(.)@samples$tissue`. If tissue is `Cell`,
then details about the cell line are available in
`experimentData(.)@samples$cellLine`.

#### PMID
Documented in `pubMedIds(.)`.

#### Spatial proteomics experiment annotation

Documented in `experimentData(.)@other`:
  - **MS** (`$MS`) type of mass spectrometry experiment: iTRAQ8,
	iTRAQ4, TMT6, LF, SC, ...
  - **Experiment** (`$spatexp`) type of spatial proteomics
	experiment: LOPIT, LOPIMS, subtractive, PCP, other, PCP-SILAC,
	...
  - **MarkerCol** (`$markers.fcol`) name of the markers feature
	data. Default is `markers`.
  - **PredictionCol** (`$prediction.fcol`) name of the localisation
	prediction feature data.

#### Example

```{r}
experimentData(dunkley2006)@samples
pubMedIds(dunkley2006)
otherInfo(experimentData(dunkley2006))

## all at once
pRolocmetadata(dunkley2006)
```

### Adding new data

The procedure to data in pRolocdata is as follows. Here, we assume
that 3 new data files are available from the manuscript of *Smith et
al. 2017*, and these files will be added to `pRolocdata` as three
`MSnSet` objects.


1. the original data (often from supplementary material) are added to
   `inst/extdata`, say `Smith_expA.csv`, `Smith_expB.csv` and
   `Smith_expC.csv` (the name should ideally be the same as the
   original files), and the files and provenance is documented in
   `inst/extdata/README`. If the data files are really big, then they
   should be compressed. If they are too big (for example don't fit on
   github or would substantially increase the size of the package),
   then we might decide not to added them, but they should still be
   documented in the README file and the script (see point 2) should
   still assume they are there.

2. A script, typically called `Smith2017.R`, is added to
   `inst/scripts/`. That script reads the files above and saves the
   corresponding (compressed) MSnSet objects directly in data,
   typically called `Smith2016a.rda`, `Smith2016a.rda`, ..., and the
   objects themselves would be named `Smith2016a`, `Smith2016b`, ...

3. Write a `man/Smith2016.Rd` documentation file documenting all
   relevant data objects, providing some information about the
   experiment and data provenance, and a reference to the original
   paper.

4. Build and check the package and, if successful, send a [github pull
   request](https://github.com/lgatto/pRolocdata).


If you do not have the `R` expertise to prepare the data, send me an
email at `lg390<AT>cam<dot>ac<dot>uk` with the source `csv` files and
appropriate metadata and I will add it for you.
