Spatial proteomics datasets
============================

## The `pRolocdata` package

`pRolocdata` is a [Bioconductor](http://bioconductor.org/)
[experiment package](http://bioconductor.org/packages/release/BiocViews.html#___ExperimentData)
([release](http://bioconductor.org/packages/release/data/experiment/html/pRolocdata.html)
and
[devel](http://bioconductor.org/packages/devel/data/experiment/html/pRolocdata.html)
pages) that collects published (mainly, although some unpublished
datasets are also available) mass spectrometry-based spatial/organelle
dataset. The data are distributed as `MSnSet` instances (see the
[`MSnbase`](http://www.bioconductor.org/packages/release/bioc/html/MSnbase.html)
for details) and are used throughout the
[`pRoloc`](http://bioconductor.org/packages/release/data/experiment/html/pRolocdata.html)
and
[`pRolocGUI`](http://bioconductor.org/packages/devel/bioc/html/pRolocGUI.html)
software for spatial proteomics data analysis and visualisation.

### Installation

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library("knitr")
library("pRolocdata")
x <- data.frame(pRolocdata()$results[, -(1:2)])
colnames(x) <- c("Data", "Description")
```

```{r, eval=FALSE}
source("http://bioconductor.org/biocLite.R")
biocLite("pRolocdata")
```

Once installed, the package needs to be loaded

```{r, eval=FALSE}
library("pRolocdata")
```

### Available datasets

Currently, there are `r nrow(x)` datasets available in
`pRolocdata`. Use the `pRolocdata()` function to obtain a list of data
names and their description.

```{r, eval=FALSE}
pRolocdata()
```

```{r, echo=FALSE, results='asis'}
kable(x, format = "markdown")
```
### Loading data

Data is loaded into the `R` session using the `load` function; for
instance, to get the data from
[Dunkley et al (2006)](http://www.pnas.org/content/103/17/6518.abstract),
one would type

```{r}
data(dunkley2006)
```

To get more information about a given dataset, see its manual page

```{r, eval=FALSE}
?dunkley2006
## or
help("dunkley2006")
```

## The datasets

Each data object in `pRolocdata` is available as an `MSnSet`
instance. The instances contain the actual quantitative data, sample
and features annotations (see `pData` and `fData`
respectively). Additional MIAPE data
[[1](https://en.wikipedia.org/wiki/Minimum_Information_About_a_Proteomics_Experiment),
[2](http://www.nature.com/nbt/journal/v25/n8/abs/nbt1329.html)]
experimental data is available in the `experimentData` slot, as
described in section *Required metadata* below.

The source of these data is generally one or several text-based
spreadsheet (`csv`, `tsv`) produced by a third-party
application. These original files are often distributed as
supplementary material to the research paper and used to generate the
`R` objects. These source spreadsheets are available in the package's
`inst/extdata` directory. The `R` script files, that read the
spreadsheets and create the `R` data is distributed in the
`inst/scripts` directory.

### Required metadata

Additional metadata is available with the `pRolocmetadata()` function
as detailed below.

#### Species
Documented in `experimentData(.)@samples$species`

#### Tissue

Documented in `experimentData(.)@samples$tissue`. If tissue is `Cell`,
then details about the cell line are available in
`experimentData(.)@samples$cellLine`.

#### PMID
Documented in `pubMedIds(.)`.

#### Spatial proteomics experiment annotation

Documented in `experimentData(.)@other`:
  - **MS** (`$MS`) type of mass spectrometry experiment: iTRAQ8,
    iTRAQ4, TMT6, LF, SC, ...
  - **Experiment** (`$spatexp`) type of spatial proteomics
    experiment: LOPIT, LOPIMS, subtractive, PCP, other, PCP-SILAC,
    ...
  - **Type** (`$type`) type of data: new, meta, both
  - **MarkerCol** (`$markers.fcol`) name of the markers feature
    data. Default is `markers`.
  - **PredictionCol** (`$prediction.fcol`) name of the localisation
    prediction feature data.

#### Example

```{r}
experimentData(dunkley2006)@samples
pubMedIds(dunkley2006)
otherInfo(experimentData(dunkley2006))

## all at once
pRolocmetadata(dunkley2006)
```

### Adding new data

New data consists of

- the source data (generally a `csv` file) in `inst/extdata` and their
  description (including origin of the files) in
  `inst/extdata/README`. The name of the source files should, if
  possible, not be modified (except for the file extension).

- the `R` script, named `newdata.R` in the `inst/scripts` directory,
  that reads the source files and generates a valid `MSnSet` instance
  (with appropriate pRolocmetadata fields) in the `data` package
  directory. The `R` data file should be names `newdata.rda`.

- A documentation file for the new data: `man/newdata.Rd`.

A [github pull request](https://github.com/lgatto/pRolocdata) with the
above can be send directly. 

Alternatively, if you do not have the `R` skills to prepare the data,
send me an email at `lg390<AT>cam<dot>ac<dot>uk` with the source `csv`
files and appropriate metadata and I will add it for you.
